---
title: "Homework 2"
author: "Zoey Zhao"
date: "10/5/2021"
output: github_document
---
```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
Read and clean the Mr. Trash Wheel sheet
```{r, message=FALSE, warning=FALSE}
trash_data <- 
  #read the excel and let it contain data entries only
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, range = "A2:N408")%>%
  #use reasonable variable names
  janitor::clean_names()%>%
  #omit rows that do not include dumpster-specific data
  drop_na(dumpster)%>%
  #round the number of sports balls to the nearest integer
  mutate(sports_balls = round(sports_balls, digits = 0))

```

Read and clean precipitation data for 2018
```{r, message=FALSE, warning=FALSE}
pre2018 <- 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14")%>%
  #omit rows without precipitation data
  drop_na(Total)%>%
  #add a variable for year
  mutate(year = 2018)

#Read and clean precipitation data for 2019
pre2019 <- 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 4, range = "A2:B14")%>%
  #omit rows without precipitation data
  drop_na(Total)%>%
  #add a variable for year
  mutate(year = 2019)

#combine precipitation datasets
pre_data <- 
  bind_rows(pre2018, pre2019)%>%
  #convert month to a character variable
  mutate(Month = month.name[Month])
```
The data set `trash_data` has `r length(trash_data)` observations including the `r names(trash_data)`. It has a total of `r nrow(trash_data)` rows and `r ncol(trash_data)` columns. `month` is a character variable (e.g. "January", "February", "March"), `year` is a numeric variable (e.g. 2017), the variable `date` looks like yyyy-mm-dd, `weight_tons` is a numeric variable accurate to two decimal places, `homes_powered` is a numeric variable accurate to five decimal places, other variables are all numeric variables of integers. The data set `pre_data` has `r length(pre_data)` observations including the `r names(pre_data)`. It has a total of `r nrow(pre_data)` rows and `r ncol(pre_data)` columns. `month` is a character variable (e.g. "January", "February", "March"), `year` is a numeric variable that only contains 2018 and 2019, and `Total` is a numeric variable accurate to two decimal places.

For available data, total precipitation in 2018 is `r sum(pull(pre2018, Total))`, and the median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trash_data, year==2019), sports_balls))`.

## Problem 2
Read and clean the data in `pols-month.csv`:
```{r, message=FALSE, warning=FALSE}
pols_month <- 
  read_csv("./data/pols-month.csv")%>%
  #break up the variable mon into variables year, month, and day
  separate(mon, into = c("year", "month", "day"), sep = "-")%>%
  #change variables year, month, and day to integers
  mutate(year = as.numeric(year))%>%
  mutate(month = as.numeric(month))%>%
  mutate(day = as.numeric(day))%>%
  #replace month number with month name
  mutate(month = month.name[month])%>%
  #create a president variable taking values gop and dem
  mutate(president = case_when(prez_gop == 1 ~ "gop",
                               prez_gop == 0 ~ "dem"))%>%
  #remove prez_dem and prez_gop; and remove the day variable
  select(-prez_gop, -prez_dem, -day)%>%
  #arrange according to year and month
  arrange(year, month)
```
Read and clean the data in `snp.csv`:
```{r, message=FALSE, warning=FALSE}
snp_data <- 
  read_csv("./data/snp.csv")%>%
  #break up the variable date into variables month, day, and year
  separate(date, into = c("month", "day", "year"), sep = "/")%>%
  #change variables month, day, and year to integers
  mutate(month = as.numeric(month))%>%
  mutate(day = as.numeric(day))%>%
  mutate(year = as.numeric(year))%>%
  #replace month number with month name
  mutate(month = month.name[as.numeric(month)])%>%
  #change year to 4 digits
  mutate(year = case_when(year > 20 ~ year + 1900,
                          year < 20 ~ year + 2000))%>%
  #arrange according to year and month
  arrange(year, month)%>%
  #remove the day variable and organize so that year and month are the leading columns
  select(year, everything(), -day)
```
Read and tidy the data in `unemployment.csv`:
```{r message=FALSE, warning=FALSE}
unemployment <- 
  read_csv("./data/unemployment.csv")%>%
  #switching from “wide” to “long” format
  pivot_longer(Jan:Dec,
               names_to = "month",
               values_to = "unemployment_percentage")%>%
  #change month from abbreviation to full name
  mutate(month = month.name[match(month, month.abb)])%>%
  #change the column name "Year" to "year"
  mutate(year = Year)%>%
  select(year, everything(), -Year)%>%
  #arrange according to year and month
  arrange(year, month)
```
Merging `snp_data` into `pols_month`, and merging `unemployment` into the result:
```{r message=FALSE, warning=FALSE}
pols_snp_unemployment <- 
  left_join(pols_month, snp_data)%>%
  left_join(unemployment)
```
The data set `pols_month` has `r nrow(pols_month)` rows and `r ncol(pols_month)` columns with column names: `r names(pols_month)`. Range of year is (`r range(pull(pols_month, year))`).  
The data set `snp_data` has `r nrow(snp_data)` rows and `r ncol(snp_data)` columns with column names: `r names(snp_data)`. Range of year is (`r range(pull(snp_data, year))`).  
The data set `unemployment` has `r nrow(unemployment)` rows and `r ncol(unemployment)` columns with column names: `r names(unemployment)`. Range of year is (`r range(pull(unemployment, year))`).  
The final resulting data set `pols_snp_unemployment` has `r nrow(pols_snp_unemployment)` rows and `r ncol(pols_snp_unemployment)` columns with column names: `r names(pols_snp_unemployment)`. Range of year is (`r range(pull(pols_snp_unemployment, year))`)

## Problem 3

